---
title: "Verbose exponence in SVCs"
author: Naomi Peck
date: today
format: 
  html:
    toc: true
    toc-expand: 2
    embed-resources: true
    smooth-scroll: true
    respect-user-color-scheme: true
editor: visual
bibliography: [references.bib, packages.bib]
theme:
  light: flatly
  dark: darkly
execute:
  echo: false
  freeze: auto
  results: "asis"
callout-icon: false
---

```{r}
#| label: set-up
#| warning: false

library(lingglosses)
library(knitr)

```

```{r}
#| label: packages-bib
#| include: false

knitr::write_bib(.packages(), file = 'packages.bib', tweak = TRUE)
```

::: {.callout-caution}
Please note that this site represents a work in progress and should not be cited without prior permission from the author.
:::

# Poster

```{r}
#| label: poster
#| out-width: 100%
#| out-height: 400

knitr::include_graphics("verbose-exponence.pdf")

```


# Introduction


-   most approaches to marking in SVCs look at the distribution of marking and create typologies based on this

    -   e.g. Aikhenvald [-@aikhenvald2006; -@aikhenvald2018]
        - single marking: category marked once per construction
        - concordant marking: category marked on every component of serial verb
            - truncated: marking is only partially the same
            - optional concordant marking: some of the markers can be omitted without change in the semantics of the construction
    -   e.g. @ross2021:
        - isolating (or indeterminate): no marking present for a category
        - sharing: one verb with inflection for a given category, other verb(s) unmarked but share features
        - agreeing: each verb marked with parallel inflection for a given category
        
-   we can move to a more predictive approach by leveraging the typology of verbose exponence [@carroll2022]

    -   why?
        -   moving to a comprehension-centric approach better captures what we think about language use and language change [cf. @carrollbeniamine2025]
    -   how?
        -   firstly, we need to focus on the domain of a feature, rather than the specific pattern of exponence^[Note: most discussions of verbose exponence limit their discussion to the word level following @harris2017. Here, I work with the construction as the relevant unit, as this level also excludes processes on the sentence level [cf. also @caballeroharris2012; @fenger2023].]
        -   different patterns of marking are linked to different paradigms of SVCs in a language
            -   and these paradigms involve different levels of informativeness associated with marking possibilities
                -   and therefore different patterns of redundancy
            -   and these patterns are crucially language-specific

-   we can use insights from interactional linguistics to generate predictions about how exponence is leveraged in real-time use

    -   in interaction, we "project" syntactic structures and prosodic structures [@auer2024]
        -   these syntactic and prosodic projects are oriented to by co-interactants
    -   the idea of "projects" is particularly useful when we have multi-part structures - like serial verb constructions
        -   what happens when only one part of the construction is realised?
            -   do speakers posit a finished or an unfinished syntactic project?
                -   is there competition between simple and complex predicates?
            -   can speakers co-construct the construction (complete the syntactic project)?
        -   what happens when speakers use afterthoughts?
            -   given its (semantic, pragmatic) compatibility, is it interpreted as an add-on to a previously finished syntactic project or a completely new one?
        -   how are repairs carried out?
            -   are syntactic projects restarted or continued?
    -   different patterns of marking with different possibilities for syntactic projection (i.e. exponence patterns) should function differently in interaction

-   taking a comprehension-based interactional perspective can help us also better understand the trajectory of SVCs

    -   the role of alignment of prosodic and syntactic projects are argued to play a crucial role in the formation of phrase structure [@himmelmann2022]
        -   how do syntactic and prosodic projects align with different exponence patterns?
            -   are some exponence patterns reliant on prosodic projection?

-   the typology of verbose exponence can thus help us predict how:

    1.  SVCs are understood and used in discourse

        → particularly with respect to on-the-fly processing and parsing
    2.  marking in SVCs are understood and used for interactive goals 

        → syntactic projects vis-á-vis prosodic projects
        → completed vs. incomplete syntactic projects
    3.  SVCs develop (and have developed) over time 

        → informativity vs. redundancy of markers

        → competition and salience of constructional forms

        → interaction of syntax and prosody

- so let’s look at how we can re-interpret the examples of SVCs that we already know through this lens

# Exponence Patterns

## Single Exponence

-   features can be marked at most once in a given domain
    
    -   two important sub-categories
        -   languages which do not allow verbose exponence (for a given feature)
        -   languages which do allow verbose exponence (for a given feature)

### Predictions

-   speakers posit a new domain for a feature upon encountering any marker
-   if no marker is encountered, the domain is ambiguous
-   construction-internal pauses are permitted if a domain is not ambiguous

    -   i.e. construction is salient
    -   normally to do with expected edge-marking

-   semantically compatible prosodic afterthoughts cannot be integrated into a SVC

    -   they will be interpreted as a separate clause

### Examples

::: {.callout-note collapse="true"}
#### Languages without verbose exponence
::: {.callout-note collapse="true"}
##### [Apinayé](https://glottolog.org/resource/languoid/id/apin1244)

> Morphologically, there is only one mode marker for the whole construction, such that all elements fall under its scope. In regard to syntax, boundary indices and the internally cohesive distribution of elements in the sentence are indicative of the internal coherence of serialization... [@cunhadeoliveira2005, 276]

(@apinaye2) @cunhadeoliveira2005, 162, Example 106d
```{r}
#| label: apinaye
gloss_example(transliteration = "Pa tẽ tẽm.",
             glosses = "RL go fall",
             free_translation = "I fell.")
```

:::
::: {.callout-note collapse="true"}
##### [Awa-Cuaiquer](https://glottolog.org/resource/languoid/id/awac1239)
  
> The Awa Pit Serial Verb construction thus consists of two active verbs with identical subcategorization frames, which are obligatorily adjacent and occur in a temporally iconic order. The normal range of morphological and syntactic possibilities is open to the clause, with any inflection occurring on the second verb stem. The first verb stem is either bare (if consonant-final) or suffixed with *t* (if vowel-final). Any modification, such as negation, applies semantically to both verbs, but is indicated formally on the second verb. [@curnow1997, 305]

(@awa2) @curnow1997, 306, example 776
```{r}
#| label: awa-1
gloss_example(transliteration = "pa azh kway-zi",
             glosses = "sun shine drop-NONLOCUT",
             free_translation = "The sun shone out.")
```

(@awa3) @curnow1997, 306, example 775
```{r}
#| label: awa-2
gloss_example(transliteration = "Demetrio=na I-t kway-zi",
             glosses = "Demetrio=TOP go-SV drop-NONLOCUT",
             free_translation = "Demetrio left.")
```

:::
::: {.callout-note collapse="true"}
##### [Belep](https://glottolog.org/resource/languoid/id/nyal1254)

> The elements of the verb group always occur in a fixed order (they are an instance of templatic morphology; c.f. Bickel & Nichols 2007) and cannot undergo any syntactic operations except as a unit. Most elements in a verb group are not obligatory, but when they occur they have a conventionalized meaning and occur in an obligatory order. [@mccracken2012, 170]

> Intonation breaks, pauses, and repairs during the production of the verb group are most likely to occur immediately after the preface; however, these should be considered indicators of a phonological, rather than a grammatical, word boundary (Dixon and Aikhenvald 2002:24). [@mccracken2012, 330]

(@belep1) @mccracken2012, 371, example 149
```{r}
#| label: belep
gloss_example(transliteration = "ava=jara bae pê",
             glosses = "1PL.EXCL.SBJ=want bite bread",
             free_translation = "We want to eat the bread.",
             comment = "Ava jaar bae pê.")
```

:::
::: {.callout-note collapse="true"}
##### [Hakhun Tangsa](https://glottolog.org/resource/languoid/id/noct1238)

> When we have multiple verbs in a row it is hard to tell whether we have multiple predications, and if so, where one ends and the other begins. ... There are a few formal means that help us delimit a predication. This involves morphological elements marking the boundary of the verb stems on both sides in two constructions – the auxiliary construction and the prohibitive construction. In addition, the position of certain verbs and knowledge of the possible semantic relations between the verbs can help us infer whether there are multiple predications and where the boundary is.  [@boro2017, 278]

(@hakhun1) @boro2017, 276, example 511
```{r}
#| label: hakhun
gloss_example(transliteration = "vàn nɤ́ twè heʔ keʔ kuʔ t-iʔ",
             glosses = "fire LOC take keep go give PST-1PL",
             free_translation = "(We) throw (it) in the fire.")
```

:::
::: {.callout-note collapse="true"}
##### [Kalam](https://glottolog.org/resource/languoid/id/kala1397)

> Kalam, a Papuan (non-Austronesian) language of the New Guinea Highlands, allows constructions in which one or more uninflected verb stems precede an inflected verb stem. Such constructions correspond to the cross-linguistically attested phenomenon of serial verb constructions (SVCs). [@lane2007, 1]

> It seems likely that the ease with which long stretches of material can intervene between SVC-initial motion stems and the following stem contributes to the possibility of ‘getting lost’ in this way. It is also possible for a speaker to pause when *am-* or *ap-* are followed by long NPs, either immediately after the verb stem itself or after the following NP. In contrast, there would be no pause in an SVC such as *pak dad apl* ‘hit carrying having-come’. The intervention of long stretches of non-verb material, and the concomitant possibility of a pause and of the recapitulation of the subject, are evidence of a weaker bond between motion verbs and the following stem(s) than the bond which usually holds between the stems of an SVC. [@lane2007, 83]


(@kalam1) @lane2007, 89, example 4.40
```{r}
#| label: kalam
gloss_example(transliteration = "...kti am kmn pak dad ap-l nb okok...",
             glosses = "they go game_mammal hit carry come-SS.PRIOR place these",
             free_translation = "... they having gone (and) hit and carried back game mammals...")
```
    
    
:::
::: {.callout-note collapse="true"}
##### [Kéo](https://glottolog.org/resource/languoid/id/keoo1238)

> The following criteria are diagnostic of serial verb constructions in Kéo: <br>
1. A serial verb construction contains two or more verbs, which occur in the same clause describing a single event; <br> 
2. The whole serial verb complex is negated by a negator preceding the first verb (see 12.6.1);<br> 
3. All verbs come under the scope of one aspect or mood particle (see 12.2 and 12.3) which precedes the verb complex;<br> 
4. All verbs share a single subject, at the surface level;<br> 
5. The construction falls under the one intonation contour. That is, there are no pauses between the verbs in a serial construction. [@baird2002, 286-287]

> Based on auditory analysis, the verb complex falls under the one intonation contour. Therefore, in fluent speech there are no pauses with a grammatical function between contiguous verbs in serial verb constructions.<br>
Intonation does not fall nor rise on non-final verbs in a serial verb complex. A break in intonation, either in the form of a marked rise or fall on a non-final verb in a series, indicates that the verbs belong to different clauses. [@baird2002, 291]

(@keo1) @baird2002, 292, example 11
```{r}
#| label: keo-1
gloss_example(transliteration = "'Imu kai mbana pasa rédé So'a.",
             glosses = "3SG go walk market up So'a",
             free_translation = "She went to the market in So'a.")
```

(@keo2) @baird2002, 292, example 12
```{r}
#| label: keo-2
gloss_example(transliteration = "'Imu kai, mbana pasa rédé So'a.",
             glosses = "3SG go walk market up So'a",
             free_translation = "She went, (she) went (walking) to the market in So'a.")
```
:::
::: {.callout-note collapse="true"}
##### [Mandarin Chinese](https://glottolog.org/resource/languoid/id/mand1415)

> Sentence \ref(@mandarin2) will not be taken to mean (2) my writing of a letter on his behalf (‘I am writing a letter for him’) unless realized with a intonational break signaling an afterthought construction (§3.4): *Wó xiě xìn, gěi ta.* ‘I am writing a letter; for him, that is.’ But like \ref(@mandarin1), this special prosodic pattern also allows for reading (1) ‘I am writing a letter; to him, that is.’. [@wiedenhof2015a, 137]

(@mandarin1) @wiedenhof2015a, 136, example 5.34a
```{r}
#| label: mandarin-1
gloss_example(transliteration = "wó gěi ta xiě xìn",
             glosses = "1.SG give 3 write letter",
             free_translation = "literally ‘I bestow on him and write a letter’",
             comment = "(i) ‘I am writing him a letter’ (the letter being written to him)
             (ii) ‘I am writing a letter for him’")
```

(@mandarin2) @wiedenhof2015a, 136, example 5.34b
```{r}
#| label: mandarin-2
gloss_example(transliteration = "wó xiě xìn gěi ta",
             glosses = "1.SG write letter give 3",
             free_translation = "literally ‘I write a letter and bestow on him’ >",
             comment = "‘I am writing a letter to him’")
```
:::
::: {.callout-note collapse="true"}
##### [Mangghuer](https://glottolog.org/resource/languoid/id/minh1238)

> In a main verb + auxiliary construction, the main verb is nonfinite. There are six morphological possibilities. In many instances, the main verb appears as just a verb stem, with no morphology at all. ...
An auxiliary verb always follows the main verb. The auxiliary verb (the final one, if more than one is present in a particular clause) bears finite morphology, or, if the clause is dependent, it is the auxiliary verb which bears whatever nonfinite morphology is used to define the relationship between the clause and another clause, on which it is dependent.[@slater2005, 132]

> No such postposing of zero-marked non-final clauses is possible. These clauses rely on their position (preceding another clause) to give them a semantic interpretation (preceding the action of the other clause in time). The fact that zero-marked clauses cannot be postposed thus appears to be not a syntactic fact, but rather, simply a semantic fact. If the zero marker had more semantic content, as do some of the other markers we have seen, zero-marked clauses might also be postposable. However, due to the lack of semantic specificity which obtains when no suffix at all is used, linear positioning must bear an unusual semantic load, and postposing becomes impossible. [@slater2005, 262]

(@mangghuer1) @slater2005, 135, example 43
```{r}
#| label: mangghuer-1
gloss_example(transliteration = "nughuai pusa di ge-jiang.",
             glosses = "dog another eat do-OBJ:PERF",
             free_translation = "the dog also ate (it).")
```

(@mangghuer3) @slater2005, 259, example 81
```{r}
#| label: mangghuer-2
gloss_example(transliteration = "zaohuang=du xi shuer=ni bari gui-jiang.",
             glosses = "kitchen=DAT go chopsticks=ACC take run-OBJ:PERF",
             free_translation = "(she) went into the kitchen, took the chopsticks and ran.")
```

:::
::: {.callout-note collapse="true"}
##### [Northern Yukian](https://glottolog.org/resource/languoid/id/yuki1243)

> Non-final verbs can be bare verb roots or can be verb roots affixed with derivational morphology, in Yuki serial verb constructions. The non-final verbs in the serial verb construction are never marked for tense or with *=mil*. The final verb in the sequence will be affixed with finite *=mil*. [@balodis2011, 665]

(@yuki1) @balodis2011, 666, example 83
```{r}
#| label: yuki-1
gloss_example(transliteration = "sąmi ki-mas-i hąša hąp šuˀ kop-wok-tl=mil tat-k-il",
             glosses = "and:then DST-DSTR-ANIM again sing sit feather-dance-TR=FIN fix-PNCT-MPSV",
             free_translation = "And then in turn these others sing, sit, and dance the feather dance and fix themselves up.",
             annotation = "sámi: kimáse hąšáˀ hąp šú: kopwóktlmil tá:tkí:li.")
```

(@yuki2) @balodis2011, 667, example 84
```{r}
#| label: yuki-2
gloss_example(transliteration = "siˀi lak-t nąw-k-il=mil ki ˀipsak",
             glosses = "then:HSAY go.out-INTR see-PNCT-MPSV=FIN DST boy",
             free_translation = "Then going out, the boy looked.",
             annotation = "seˀéy lákti nąwkílmil kí ˀipsák")
```
:::
::: {.callout-note collapse="true"}
##### [Uduk](https://glottolog.org/resource/languoid/id/uduk1239)

> In Uduk serial verb constructions, there are at least two verb roots in a single predicate, the first verb of the series bearing finite inflectional morphology and further verbs occurring in a non-finite form. Tonal aspectual changes and argument cross-referencing are possible only for the first verb of the construction; further verbs must stay consistent with scope and aspect. [@killian2015, 248]

(@uduk1) @killian2015, 246, example 773
```{r}
#| label: uduk-1
gloss_example(transliteration = "wàthíʔ dǒsh-óʼd ì yà ʼdéʼk shōʼk āràbíyāʔ",
             glosses = "man stand:IPFV-3SG LK go lift foot vehicle",
             free_translation = "The man is standing and goes to lift a tire.")
```

(@uduk2) @killian2015, 249, example 785
```{r}
#| label: uduk-2
gloss_example(transliteration = "gù'b rús-ù 'bár mò",
             glosses = "house mud:PFV-IMPRS finish PART",
             free_translation = "All of the house was mudded.")
```
:::
:::

::: {.callout-note collapse="true"}

#### Languages with verbose exponence
::: {.callout-note collapse="true"}
##### [Menya](https://glottolog.org/resource/languoid/id/meny1245)

> ...verb serialization in Menya typically involves the juxtaposition of two verbs, one that is stripped of all the suffixation normally associated with verbs and the other an inflected verb of motion. [@whitehead2005, 181]

(@menya1) @whitehead2005, 182, example 431
```{r}
#| label: menya
#| results: asis
gloss_example(transliteration = "ye ä-p-wä-ayi qänakä=i ä-ma-ät-a ä-ma ä-p-y-ayi",
             glosses = "1DU ASS-come-go_down-1DU.SS pandanus=DEF ASS-get-SS-1DU.SS ASS-get ASS-come-go_up-1DU.SS",
             free_translation = "...we went down, got the pandanus nuts and brought them up, then...", 
             annotation = "...ye äpäweyi, qänaki ämeta äma äpäyayi...")
```

:::
::: {.callout-note collapse="true"}
##### [Sabanê](https://glottolog.org/resource/languoid/id/saba1268)

More specifically, marking of object arguments (and also TAM to some extent):

> When an identical agent/subject is involved in two actions in coordinated sentences, the verbs can be concatenated in which case only the last one receives all verbal suffixes. ... However, if tense or arguments are distinct, each sentence has its own TAM structure. In this case, sentences are just concatenated. [@dearaujo2004, 189-190]

(@sabane1) @dearaujo2004, 129, example 8
```{r}
#| label: sabane-1
gloss_example(transliteration = "ip-a-ntal-i a-san-i-ntal-i",
             glosses = "see-VS-PRET-ASSR 3OBJ-catch-VS-PRET-ASSR",
             free_translation = "S/he saw it and caught it.")
```

(@sabane2) @dearaujo2004, 152, example 122
```{r}
#| label: sabane-2
gloss_example(transliteration = "sa-n-n t-osa-n-n",
             glosses = "catch-VS-IMP 1OBJ-give-VS-IMP",
             free_translation = "Catch (it) and give it to me.",
             comment = "[ˌsãˑtoˈsãː]")
```

(@sabane3) @dearaujo2004, 190, example 118
```{r}
#| label: sabane-3
gloss_example(transliteration = "towali wola ilul-i iney-i-dana",
             glosses = "1SBJ a_lot eat-VS drink-VS-PRS.EVID",
             free_translation = "I eat and drink a lot.")
```

:::
::: {.callout-note collapse="true"}
##### [Saramaccan](https://glottolog.org/resource/languoid/id/sara1340)

While aspect can be marked via multiple exponence, tense and argument are only marked once per construction.

(@saramaccan3) @mcwhortergood2012, 149, example 88
```{r}
#| label: saramaccan-tense
gloss_example(transliteration = "A súti dí patupátu túwɛ gó kai a wáta.",
             glosses = "3SG shoot DEF duck throw go fall LOC water",
             free_translation = "He shot the duck and it fell into the water.")
```

:::
::: {.callout-note collapse="true"}
##### [Taba (East Makian)](https://glottolog.org/resource/languoid/id/east2440)
Glottocode: east2440 <br> 
Family: Austronesian <br>
Macroarea: Papunesia

Taba is an example of a language for which Aikhenvald identifies "optional" concordant marking for subject agreement.

(@taba1) @bowden2001, 304, example 28
```{r}
#| label: taba-2
gloss_example(transliteration = "t=han t=ronda po-pe Ploili",
             glosses = "1PL.INCL=go 1PL.INCL=stroll down-ESS Peleri",
             free_translation = "We went strlling in Peleri.",
             annotation = "Than tronda pope ploili.")
```

xxx optional marking

:::
::: {.callout-note collapse="true"}
##### [Tamasheq](https://glottolog.org/resource/languoid/id/tama1365)

> A positive clause with a Shlmpf or Imprt verb can be added to another positive clause in the future construction, without repeating Future ad. In \ref(@tamasheq1), ɑ̀d occurs once at the beginning followed by the usual Shlmpf, but a second Shlmpf clause is conjoined (without another Fut particle or other overt connective). If a direct object or dative NP occurs in both clauses, it appears in pronominal form (object or dative clitic) in the second \ref(@tamasheq2). Any clitics in the add-on clause are hosted by the verb. ... In all cases, the two clauses have separate accentuation. [@heath2005, 680]

(@tamasheq1) @heath2005, 680, example 828a
```{r}
#| label: tamasheq-1
gloss_example(transliteration = "ɑ̀d n-ækš n-əsə́w",
             glosses = "FUT 1PL.SBJ-eat.IPFV 1PL.SBJ-drink.IPFV",
             free_translation = "We will eat and drink.")
```

Note that the orthography in \ref(@tamasheq2) has been altered slightly; please consult the original source for the correct rendering.

(@tamasheq2) @heath2005, 681, example 828c
```{r}
#| label: tamasheq-2
gloss_example(transliteration = "ɑd ə̀wət-Aɤ t-è-hæle æ̀ŋɤ-æq-qæt",
             glosses = "FUT hit.IPFV-1SG.SBJ FEM-SG-sheep kill.IPFV-1SG.SBJ-3SG.FEM.OBJ",
             free_translation = "I will hit and kill the sheep.")
```
:::
::: {.callout-note collapse="true"}
##### [Tondano](https://glottolog.org/resource/languoid/id/tond1251)

> The broad definition of an SVC applied here follows that of Haspelmath (ibid: 2) which states that an SVC is “a monoclausal construction consisting of multiple independent verbs with no element linking them and with no predicate-argument relation between the verbs”. <br>
In addition to this overarching feature, SVCs in Tondano also have the following characteristics (these are also considered characteristics of SVCs in Durie 1997:291, Aikhenvald 2006, Dixon 2006, and Noonan 2007:65, 88): <br>
- SVCs are comprised of multiple verbs which occur together within a clause. Other elements (e.g. conjunctions, negators, NPs, or PPs) do not occur between the two verbs. <br> 
- The multiple verbs function as the head of a single predicate, and more importantly, they refer to a situation which is viewed as a single overall event. <br>
- Each verb in the SVC may also function independently as the sole verb in a verbal clause. <br>
- An SVC has only one shared pivot argument. [@brickell2014, 424]

> Example \ref(@tondano2) again displays two separate clauses, despite the absence of a co-ordinating conjunction. The reason this can only be an instance of parataxis is due to a number of features, these are: the pause in intonation between the two verbs, the difference in TAM marking (*<um>* is irrealis while *meN-* is realis), the separate representations of the PIV arguments, and the fact that there are two separate assertions (events) denoted by the verbal predicates.


(@tondano1) @brickell2014, 425, example 772
```{r}
#| label: tondano-1
gloss_example(transliteration = "ko=<um>èa s<um>iwo litir mana=mèè",
             glosses = "2SG.PIV=<AV>go <AV>make dike there=DIR.MED",
             free_translation = "Will you go and build dikes over there?",
             annotation = "komèa sumiwo litir manamèè?")
```

(@tondano2) @brickell2014, 427, example 776
```{r}
#| label: tondano-2
#| results: asis
gloss_example(transliteration = "sèa {<um>upu=mow} sè=meN-CVCV-kantar",
             glosses = "3SG {<AV>pick=CPL} 3PL.PIV=AV.DYN-RDP-sing",
             free_translation = "They would pick (rice), they are singing.",
             annotation = "sèa mupumou, èmengantangantar")
```
:::
:::

## Multiple Exponence

-   the same feature is marked multiple times in the same domain [cf. @harris2017]

    - "fully superfluous exponence" [@caballeroharris2012; @harris2017; @grofulovicmuller2023]
    -   marker identity
    -   marker non-identity

### Predictions

-   upon encountering a marker with the same value, speakers understand this as belonging to the same domain
-   if a different value is encountered, a new domain is posited
-   construction-internal pauses are not permitted when markers are the same

    -   i.e. SVCs look like the juxtaposition of two finite clauses, and as such, a verb which can be understood as a standalone clause before a pause is analysed as such
-   compatible prosodic afterthoughts can be integrated into a SVC?

    -   although it may be ambiguous whether they constitute a separate clause
    -   must be semantically/prosodically plausible according to patterns in the language

### Examples

::: {.callout-note collapse="true"}

#### Same Marker
::: {.callout-note collapse="true"}
##### [Ashéninka Perené](https://glottolog.org/resource/languoid/id/ashe1272)

Each verb is marked separately for person, aspect, realis status, and mode. The two verbs must be realised without an intervening pause [@mihas2015, 163].

(@asheninka1) @mihas2015, 165, example 6.13.b
```{r}
#| label: asheninka-1
gloss_example(transliteration = "o-ja-t-atz-i o-shimaa-t-a",
             glosses = "3NM.S-go-EP-PROG-REAL 3NM.S-fish-EP-REAL",
             free_translation = "She went to fish.")
```

:::
::: {.callout-note collapse="true"}
##### [Bardi](https://glottolog.org/resource/languoid/id/bard1255)

> [Serial verbs] occur in a single intonation contour (although there are also examples with breaks, and examples where intonation units and syntactic units are not isomorphic). [@bowern2013, 663-664]

> A further piece of evidence shows that these clauses are joined in a single higher structure rather than chained without embedding. It is possible to extract constituents from a lower clause into a higher clause. If the clauses are conjoined, this should not be possible, since conjunction is a strong barrier to extraction... [@bowern2013, 664-665]

(@bardi1) @bowern2013, 663, example 16.64
```{r}
#| label: bardi
#| results: asis
gloss_example(transliteration = "Ginyinggon roowil i-n-nya-na Ngarrigoonbooroo baali-ngan darr i-n-ar-na=jirri niimana aamba agal ambooriny Ngoolbirndi.",
             glosses = "then walk 3-TR-catch-REM.PST {NAME} shade-ALL come 3-TR-spear-REM.PST=3AUG.IO many men and people {NAME}",
             free_translation = "Then Ngarrigoonbooroo walked to her camp and came across many people at Ngoolbirndi.",
             annotation = "Ginyinggon roowil innyana Ngarrigoonbooroo baalingan darr inarnajirri niimana aamba agal ambooriny Ngoolbirndi.")
```

:::
::: {.callout-note collapse="true"}
##### [East Ambae](https://glottolog.org/resource/languoid/id/east2443)

> Intonation is an important factor in recognising SVCs, and in distinguishing them from conjoined clauses. While in many cases the only possible reading for a particular clause requires that it be recognised as containing an SVC, in numerous other cases ambiguity could arise, if it were not for the fact that the clause boundaries were indicated by intonation. Sentences \ref(@eambae1) and \ref(@eambae2) are minimal pairs which can only be distinguished by the difference in intonation. The comma in these sentences indicates the clause boundary. ... Sentence \ref(@eambae1) contains two clauses, the first of which is an example of aspectual serialisation[,] where *rovo* indicates completive aspect. Sentence \ref(@eambae2) is not an example of an SVC, but here *rovo* has a discourse function in linking clauses, expressing the fact that 'after' the event ofthe first clause, another event occurred. [@hyslop2001, 275]

(@eambae1) @hyslop2001, 275, example 3
```{r}
#| label: eambae-1
gloss_example(transliteration = "Mo vatu na veveo mo rovo, mo vai na hinaga...",
             glosses = "RL weave ACC weaving RL finish RL make ACC food",
             free_translation = "She finished weaving the weaving, and she made the food...")
```

(@eambae2) @hyslop2001, 275, example 4
```{r}
#| label: eambae-2
gloss_example(transliteration = "Mo vatu na veveo, mo rovo, mo vai na hinaga...",
             glosses = "RL weave ACC weaving RL finish RL make ACC food",
             free_translation = "She did the weaving, and after that she made the food...")
```

:::
::: {.callout-note collapse="true"}
##### [Hatam](https://glottolog.org/resource/languoid/id/hata1243)

> Hatam allows sequences of verbs without intervening conjunctions, usually called serial verb constructions. Although there are instances of three (or more) verbs, I will make my comments on sequences of two verbs. In most cases both verbs are marked for subject person/number... [@reesink1999, 97]

> When explicit conjunctions are employed it is clear that the sequence involves coordination of clauses. The same observation holds for sequences interrupted by a pause, as in example \ref(@hatam2). [@reesink1999, 98]

(@hatam1) @reesink1999, 97, example 4.95
```{r}
#| label: hatam-1
gloss_example(transliteration = "Munggwom de di-ma i-pim i-nggum lo ig-yem hum ngum.",
             glosses = "child POS REL-that 3PL-cry 3PL-hungry but 3PL-eat fire coal",
             free_translation = "Those children are crying from hunger but they are eating charcoal.")
```

(@hatam2) @reesink1999, 97, example 4.97
```{r}
#| label: hatam-2
gloss_example(transliteration = "Lene i-ngot igy-o, i-com tut paku ni-kinei-o...",
             glosses = "then 3PL-tie house-or 3PL-throw along nail 3SG-bad-or",
             free_translation = "Then they built houses, they threw away nails that were bad...")
```
:::
::: {.callout-note collapse="true"}
##### [Kayardild](https://glottolog.org/resource/languoid/id/kaya1319)

In particular, ACT:

> In (@kayardild2) the verbs belong to different clauses, as shown by the independent intonation contours (indicated here by a comma); a literal meaning is forced. ... In (@kayardild3) ... *thaatha* is a clausemate and follows the main verb, giving the 'go and return' meaning. [@evans1995, 308]

(@kayardild2) @evans1995, 308, example 8-45
```{r}
#| label: kayardild-2
gloss_example(transliteration = "niya kuujuu-j, thaa-th",
             glosses = "3SG.NOM swim-ACT return-ACT",
             free_translation = "He had a swim, and came back.")
```

(@kayardild3) @evans1995, 308, example 8-47
```{r}
#| label: kayardild-3
gloss_example(transliteration = "niya kuujuu-ja thaa-th",
             glosses = "3SG.NOM swim-ACT return-ACT",
             free_translation = "He's gone for a swim (and will come back).")
```

:::
::: {.callout-note collapse="true"}
##### [Ket](https://glottolog.org/resource/languoid/id/kett1243)

> Ket verbs can stand on their own as independent sentences. Therefore, it is difficult to know whether two clauses are combined into a complex construction or rather constitute two standalone sentences linked in discourse. The criterion of intonation does not really seem to be of much help here. [@nefedov2015, 95]

(@ket1) @nefedov2015, 107, example 4.51
```{r}
#| label: ket
gloss_example(transliteration = "dɨˀl du-k-hən du-den",
             glosses = "child 3-TH-stand 3-weep",
             free_translation = "The child is standing (and) crying.",
             annotation = "dɨˀlʲ dúkkʌːn dúrɛn")
```

:::
::: {.callout-note collapse="true"}
##### [Lavukaleve](https://glottolog.org/resource/languoid/id/lavu1241)

In particular, object marking:

> Serial verb constructions consist of two or, rarely, three verbs in juxtaposition, functioning as a single predicate, with no morphology to mark their relationship with each other. Each of the verbs receives its own stress, and thus is a separate phonological word, but the construction as a whole is expressed in one intonational unit. Verbs of a serial verb construction share their subjects and (if more than one member is transitive) objects. They also share adjuncts. They share all their morphology except object marking; all transitive verbs cross-reference their own objects. However, all other categories that can be marked on verbs (e.g. subject marking, TAM, polarity, valency changing, and other morphology) are marked only once on the predicate, and have scope over the whole predicate. [@terrill1999, 339-340]

(@lavukaleve3) @terrill1999, 207, example 4
```{r}
#| label: lavukaleve
#| results: asis
gloss_example(transliteration = "lo fi o-ha-n va vo-fou vo-a-kuru-re filifilimal vo-na",
             glosses = "finish 3SG.NEUT.FOC 3SG.POSS-rib_of_canoe-PL ART.PL 3SG.PL.OBJ-put_on 3SG.PL.OBJ-1SG.SBJ-hit-FUT nails 3PL.OBJ-in",
             free_translation = "Having finished, I make the ribs and hammer them. With nails.",
             annotation = "Lo fi, ohan va vofou, voakurure. Filifilimal vona.")
```

:::
::: {.callout-note collapse="true"}
##### [Maybrat-Karon](https://glottolog.org/resource/languoid/id/maib1239)

> ...[serialising] constructions seem similar to two types of construction discussed before: they are similar to *coordinating constructions* from the point of view of intonation, and morphology, and they behave largely the same when placed in the interrogative mood. However, these forms violate the CSC. [@dol2007, 220]

(@maybrat1) @dol2007, 213, example 124a
```{r}
#| label: maybrat-1
gloss_example(transliteration = "m-amo m-ate ayà /",
             glosses = "3.UV-go 3.UV-go_under water (pause)",
             free_translation = "She goes and bathes.")
```

(@maybrat2) @dol2007, 213, example 124b
```{r}
#| label: maybrat-2
gloss_example(transliteration = "m-amo / m-ate ayà /",
             glosses = "3.UV-go (pause) 3.UV-go_under water (pause)",
             free_translation = "She goes, and bathes.")
```

:::
::: {.callout-note collapse="true"}
##### [Morisyen](https://glottolog.org/resource/languoid/id/mori1278)

(@morisyen1) @syea2013, 18, example xxx 
```{r}
#| label: morisyen
gloss_example(transliteration = "zot kwi diri zot manze",
             glosses = "3PL cook rice 3PL eat",
             free_translation = "They cook rice and eat")
```

:::
::: {.callout-note collapse="true"}
##### [Taba (East Makian)](https://glottolog.org/resource/languoid/id/east2440)

Taba is an example of a language for which Aikhenvald identifies "optional" concordant marking for subject agreement.

(@taba1) @bowden2001, 304, example 28
```{r}
#| label: taba
gloss_example(transliteration = "t=han t=ronda po-pe Ploili",
             glosses = "1PL.INCL=go 1PL.INCL=stroll down-ESS Peleri",
             free_translation = "We went strlling in Peleri.",
             annotation = "Than tronda pope ploili.")
```

xxx optional marking

:::
::: {.callout-note collapse="true"}
##### [Tariana](https://glottolog.org/resource/languoid/id/tari1256)

> All the components of a serial construction are separate phonological words, since they have independent stress. They have the prosodic properties of a monoverbal clause (cf. Givón 1991). A characteristic property of a prosodic phrase in the Arawak languages of the Upper Rio Negro is the "nasal pause" marker — a phrase final syllable *-hã* — used as a boundary signal (Aikhenvald 1996b). This marker cannot be inserted between the components of serial constructions; it can only appear on the last component. In \ref(@tariana2), *-hã* 'pause marker' appears on the last component of a serial construction marking a boundary between two predicates which belong to adjacent clauses. <br>
... The insertion of the boundary marker between the two components of a serial construction would be ungrammatical (?**haɾia di-swa-hã di-a* (dawn 3SGNF-stay-PAUS 3SGNF-go)). When predicates or clauses are coordinated, there is a characteristic "enumerating" intonation. (The phrase-final syllable is lengthened and the pitch of the phrase final syllable falls.) This intonation pattern is not used with components of a serial construction. [@aikhenvald1999, 477-478]

(@tariana2) @aikhenvald1999, 477, example 7
```{r}
#| label: tariana
gloss_example(transliteration = "ne-pidana haɾia di-swa di-a-hã haria-ka di-swa",
             glosses = "then-REM.P.INFR dawn 3SGNF-stay 3SGNF-go-PAUS dawn-SUB 3SGNF-stay",
             free_translation = "Then it was dawning (PAUSE). When it had dawned (she got off a tree)...")
```

:::
::: {.callout-note collapse="true"}
##### [Tidore](https://glottolog.org/resource/languoid/id/tido1248)

> In this section it will be shown that the number of true serial verb constructions is limited indeed. In most cases, constructions that now resemble serial verb constructions are either the result of the loss of inflectional morphology, in particular the loss of the object prefix, or they are grammaticalised verbal elements that still show remnants of an erstwhile status as main predicates. [@vanstaden2000, 308]

> The distinction between parataxis and serialisation is not always straightforward. In general the ‘tightly-knit’ paratactic constructions will be closest to serialisation. Neither parataxis nor serialisation use conjunctions, and the intonation of both is often like a single clause. In the case of parataxis, however, pauses and a non-final contour on the first clause are always possible, without significantly changing the interpretation, or the interpretation possibilities of the construction... [@vanstaden2000, 280]

xxx check 280 for examples

(@tidore1) @vanstaden2000, 309, example 169
```{r}
#| label: tidore
gloss_example(transliteration = "una wo-maleko wo-reke",
             glosses = "3SG.MASC 3SG.MASC.A-continually 3SG.MASC.A-cry",
             free_translation = "He cried on and on")
```
:::
::: {.callout-note collapse="true"}
##### [Turkish](https://glottolog.org/resource/languoid/id/nucl1301)

> There are verbal sequences with certain aspectual "light" verbs that yield verbal compounds, but which are restricted to occurring with the same tense / aspect suffixes on both parts of the compound. [@kornfilt1997, 478]

(@turkish1) @kornfilt1997, 478, example 1719
```{r}
#| label: turkish-1
gloss_example(transliteration = "söyle-n-ir dur-ur",
             glosses = "tell-REFL-AOR stay-AOR",
             free_translation = "She keeps on grumbling.")
```

(@turkish2) @kornfilt1997, 478, example 1720
```{r}
#| label: turkish-2
gloss_example(transliteration = "söyle-n-di dur-di",
             glosses = "tell-REFL-PST stay-PST",
             free_translation = "She keeps on grumbling.")
```
:::
::: {.callout-note collapse="true"}
##### [Yongning Na](https://glottolog.org/resource/languoid/id/yong1270)

> Most concatenations of verbs in Na fall into one of the following categories: verb phrases which belong to separate clauses but appear adjacently because their nominal arguments are not overt; a complement taking predicate with an adjacent infinitive-like verb; a verb with an aspect marker which has grammaticalized from a verb and thus has an isomorphic verb counterpart; or a resultative/cause-and-effect SVC. ... Symmetrical serial verb constructions tend to match Aktionsart marking on each verb. [@lidz2010, 396]

(@yongningna1) @lidz2010, 397, example 566
```{r}
#| label: yongning
#| results: asis
gloss_example(transliteration = "tʰi13 lə33-wæ33 zɔ33 tʰɯ33 ʐɯ33mi33 tʰɯ33 lɯ33 lə33-wɔ13 lə33-sɩ31 lə33-yɔ33 hɯ33",
             glosses = "so ACCOMP-call CSM this hearth.room this CLS ACCOMP-return ACCOMP-live ACCOMP-come FUT.REM",
             free_translation = "Having called it, the hearth room will come back to life.")
```
:::
:::

::: {.callout-note collapse="true"}

#### Different Marker

cf. Aikhenvald's "truncated" concordant marking

::: {.callout-note collapse="true"}
##### [Avatime](https://glottolog.org/resource/languoid/id/avat1244)

(@avatime2) @defina2016, 673, example 74a
```{r}
#| label: avatime
gloss_example(transliteration = "ma-tsà tomatoes=ye a-kpɛ mí kè-zi=a mɛ̀",
             glosses = "1SG.PFV-cut tomatoes=DEF SV.1SG.PFV-put LOC C6S-bowl=DEF inside",
             free_translation = "I cut tomatoes and put them in the bowl.")
```

:::
::: {.callout-note collapse="true"}
##### [Daakaka](https://glottolog.org/resource/languoid/id/daka1243)

> One fundamental criterion according to which all serial predicate constructions can be classified into one of two groups is finite marking: Depending on the non-initial predicate, some [serial predicate constructions] contain only one TAM marker (single marking), while in others, each predicate is preceded by its own TAM marker (multiple marking). Subject pronouns are never allowed in a non-initial predicate, the TAM marker always takes its monosyllabic form, as with third person singular subjects... [@vonprince2012, 304]

(@daakaka1) @vonprince2012, 308, example 6c
```{r}
#| label: daakaka-1
gloss_example(transliteration = "mwe tavya te polo usili ló-ó swa vyan milye",
             glosses = "RL get_up CONJ climb follow plant-coconut one go on_top",
             free_translation = "she got up and climbed up a coconut palm")
```

(@daakaka2) @vonprince2012, 336, example 80a
```{r}
#| label: daakaka-2
gloss_example(transliteration = "Ko=m en seaa ada mubuo mo nok?",
             glosses = "2SG=RL eat all CL2:1DU.INCL.POSS meat RL finish",
             free_translation = "Have you eaten all our meat?")
```

:::
::: {.callout-note collapse="true"}
##### [Goemai](https://glottolog.org/resource/languoid/id/goem1240)
Glottocode: goem1240 <br> 
Family: Afro-Asiatic <br>
Macroarea: Africa

> As repeatedly stated in the literature, SVCs should not contain any marker of co- or subordination. ... Compare examples \ref(@goemai1) and \ref(@goemai2): the conjoined structure in \ref(@goemai2) can only be recognized by the presence of an intonation break. [@hellwig2006, 91]

(@goemai1) @hellwig2006, 91, example 2a
```{r}
#| label: goemai-1
gloss_example(transliteration = "sai su ru n-goede gado",
             glosses = "then run.SG enter.SG LOC-bottom bed",
             free_translation = "Then (he) ran (and) entered under the bed.")
```

(@goemai2) @hellwig2006, 91, example 2e
```{r}
#| label: goemai-2
gloss_example(transliteration = "mûep swo / mûep rwo n-s'et",
             glosses = "3PL run.PL {} 3PL enter.PL LOC-bush",
             free_translation = "They ran, (and) they entered into the bush.")
```

xxx t'ong example

:::
::: {.callout-note collapse="true"}
##### [Murrinhpatha](https://glottolog.org/resource/languoid/id/murr1258)

(@murrinhpatha2) @blythe2009, 364, On the Flat line 103
```{r}
#| label: murrinhpatha-1
gloss_example(transliteration = "ngunu-murrk-nu=ngi",
             glosses = "1SG.S.7go.FUT-eat-FUT=1SG.S.1sit.FUT",
             free_translation = "I want to eat some.")
```

(@murrinhpatha3) @blythe2009, 378, Ningbingi line 13
```{r}
#| label: murrinhpatha-2
gloss_example(transliteration = "me-nge-dha=pirri",
             glosses = "3SG.S.say.PSTIMP-3SG.IO-PST=3SG.S.3stand.PSTIMP",
             free_translation = "He was saying to her...")
```

:::
::: {.callout-note collapse="true"}
##### [Olo](https://glottolog.org/resource/languoid/id/oloo1241)

> Olo has a construction that contains more than one verb and for which the verbs are not in a grammatical dependency relationship. The verbs take their full morphological marking irrespective of their serial position. The construction is under a single intonation contour. [@staley2007, 37]

(@olo2) @staley2007, 39, example 115a
```{r}
#| label: olo
gloss_example(transliteration = "ne kapi yali n-ila-pe n-e uf",
             glosses = "3FEM get.3PL firewood.PL 3FEM-lift-3PL 3FEM-go village",
             free_translation = "She got the firewood, lifted it and went to the village.")
```

:::
::: {.callout-note collapse="true"}
##### [Pech](https://glottolog.org/resource/languoid/id/pech1241)

> Verb + verb conjunction is accomplished by the use of serial verb-stems which include, for each verb involved, the stem and the subject-suffix (except for the 1st-person plural, for which only the stem is used for all but the last verb-stem), plus any associated object-prefixes... [@holt1999, 68]

(@pech1) @holt1999, 68 (glossing NP)
```{r}
#| label: pech-1
gloss_example(transliteration = "mángoh kúh-ẽ̀ʔ-pẽ́ʔ",
             glosses = "mango buy-eat-FUT:EMPH",
             free_translation = "He will buy and eat a mango.")
```

(@pech2) @holt1999, 68 (glossing NP)
```{r}
#| label: pech-2
gloss_example(transliteration = "mikátes kúha-pèʔka-uši",
             glosses = "rope buy:1SG-bring:1SG-1SG.RECPST",
             free_translation = "I bought a rope and brought it (home).")
```
:::
::: {.callout-note collapse="true"}
##### [Polish](https://glottolog.org/resource/languoid/id/poli1269)

> The evidence demonstrates that the WZIĄĆ gram can be regarded as a SVC, relatively close to the cross-linguistic prototype postulated by Aikhenvald and Dixon (2006) and Bisang (2009). To be exact, the gram complies with the definitional characteristics of the prototype of a SVC – in each such case, it does so to a high extent: <br>
6) The WZIĄĆ gram has a unitary TAM value and, consequently, the individual components cannot select TAM categories separately.<br>
...
8) Operators of time, space, manner, and instrument extend over the entire construction rather than individual components. The use of such operators for the two verbs separately is ungrammatical.<br>
9) The components of the WZIĄĆ gram invariably and compulsorily share the subject argument. More generally, the WZIĄĆ gram disallows duplicate roles, be they core and peripheral arguments, or adjuncts. The total argument structure is never more complex than the argument structure of V2. ... [@andrason2018a, 618-619]

> The WZIĄĆ gram fully complies with the phonological property of SVCs – namely, a fully audible pause comparable to a break that separates coordinated clauses (both syndetically and asyndetically) is absent in the WZIĄĆ gram (Góralczyk 2010: 100, 103; Gębka-Wolak 2012: 54). For example, in \ref(@polish1), V1 and V2 are pronounced with no pause but rather as a single phonological phrase. In fact, the pause is absent even if some elements intervene between V1 and V2. [@andrason2018a, 587-588]

(@polish1) @andrason2018a, 588, example 9a
```{r}
#| label: polish
gloss_example(transliteration = "Weź *[,] skończ ględzić",
             glosses = "take.PF.2SG.IMPR PAUSE stop.PF.2SG.IMPR complain",
             free_translation = "Stop complaining!")
```

:::
::: {.callout-note collapse="true"}
##### [Tommo So Dogon](https://glottolog.org/resource/languoid/id/tomm1242)

(@dogon1) @mcpherson2013, 442, example 895b
```{r}
#| label: dogon
gloss_example(transliteration = "Bíku=gɛ gòò-nd-áa ú=ɲ̀ òb-ì-m",
             glosses = "pen=DEF go_out-FACT-PFV 2SG.PRO=OBJ give-PFV.L-1SG",
             free_translation = "I took out the pen and gave it to you.")
```
:::
:::
## Distributed Exponence

-   multiple markers with different values must be present to encode a given feature

    -   feature realised by combination of morphemes
    -   also including here "discontinuous exponence"

### Predictions

-   upon encountering a compatible marker, speakers analyse the verb as belonging to the same domain

    -   i.e. constructional paradigm is salient
-   if an incompatible marker is encountered, a new domain is posited
-   construction-internal pauses are permitted if exponence has a salient form

    -   i.e. construction is salient to the extent that it is clear that one part of the construction is missing
-   compatible prosodic afterthoughts can be integrated into a SVC

    -   this is probably more the case if there is no salient exponence
    -   are these two things mutually exclusive?

### Examples

::: {.callout-note collapse="true"}
##### [Filomeno Mata Totonac](https://glottolog.org/resource/languoid/id/filo1235)

Specifically, TAM marking:

> When a complement-taking verb is followed by another verb with the same subject as its complement, incomplete sharing of inflection occurs. ... In such constructions, V1 is fully inflected as normally for person agreement and TAM. Whatever the aspect of V1, V2 is in the imperfective aspect, that is, it has final stress and carries the IMPF suffix –aa where expected (whenever another suffix or enclitic attaches outside it, mostly with first plural or second subjects or objects). [@mcfarland2009, 200]

(@ftotonac1) @mcfarland2009, 201, example 252
```{r}
#| label: ftotonac-1
gloss_example(transliteration = "na-k-ta-yaa-aa na-k-malakčawa-aa",
             glosses = "FUT-1SBJ-INC-stand-IPFV FUT-1SBJ-open-IPFV",
             free_translation = "I will get up to close the door.",
             annotation = "naktayáa nakmalakčawá")
```

(@ftotonac2) @mcfarland2009, xxx
```{r}
#| label: ftotonac-2
gloss_example(transliteration = "ka-tan-ti tantlii-aa-[cg]",
             glosses = "IRR-come2-2SG.SBJ dance-IPFV-2SG.SBJ",
             free_translation = "Come dance!",
             annotation = "katát tantlíiy'a")
```
xxx voiceless diacritic on i of -ti
:::

::: {.callout-note collapse="true"}
#### [Mali](https://glottolog.org/resource/languoid/id/mali1284)

Tense marking in Mali SVCs:

> The term ‘clause chaining’ is also applicable to a specific sentence level syntactic construction in Mali in which strings of short clauses are marked with a distinctive tense pattern on the final verb. ... The final verb in the chain is always in the present tense... [@stebbins2011, 291-292]

(@mali2) @stebbins2011, 292, example 18.24
```{r}
#| label: mali
#| results: asis
gloss_example(transliteration = "da ta the:k.ne:=athe:vama chases-ta va ta thet ti ngim nani kama a=ru-ta",
             glosses = "and 3H.PL send.NPRES=3H.PL.POSS spy-H.PL PURP 3H.PL go.FUT 3H.PL look.PRES for ART SPEC=person-H.PL",
             free_translation = "And they sent their spies... to go out and look for people...",
             annotation = "De: tha the:k.nathe:vama chasesta... va tha thet ti ngim nani chama arutha...")
```
:::

::: {.callout-note collapse="true"}
#### [Paamese](https://glottolog.org/resource/languoid/id/paam1238)

Specifically, mood and negative polarity:

(@paamese1) @crowley1987, 45, example 16 (glossing adjusted)
```{r}
#| label: paamese-1
gloss_example(transliteration = "kaile a-ro-muasi-tei vuasi voo-mate",
             glosses = "3PL 3PL-NEG-hit.RL-NEG pig IMM-die.RL",
             free_translation = "They didn't kill the pig by hitting it.",
             annotation = "Kail aromuastei vuas voomat")
```

(@paamese2) @crowley1987, 46, example 20 (glossing adjusted)
```{r}
#| label: paamese-2
gloss_example(transliteration = "mu-na-sooni-tei auve na-hina",
             glosses = "2PL-POT-throw.IRR-NEG throwing_stick POT-go_up.IRR",
             free_translation = "Don't throw throwing sticks up.",
             annotation = "Munasoodei auv nahina")
```

(@paamese3) @crowley1987, 47, example 23 (glossing adjusted)
```{r}
#| label: paamese-3
gloss_example(transliteration = "kaile a-muasi vuasi roo-mate-tei",
             glosses = "3PL 3PL-hit.RL pig NEG-die.RL-NEG",
             free_translation = "They hit the pig but it didn't die.",
             annotation = "Kail amuas vuas roomattei")
```

But also perhaps "ambient" serialisation?

(@paamese5) @crowley1987, 47, example 23 (glossing adjusted)
```{r}
#| label: paamese-ambient
gloss_example(transliteration = "na-mutii-nau mule eni kau-ku",
             glosses = "1SG.RL-punch.RL-1SG exist(.3SG.RL) SPEC side-1SG",
             free_translation = "I hit myself in the side.",
             annotation = "Namutiinau mul en kauk")
```
:::

## Subset Exponence

-   two important subtypes

    -   Value Subset
        -   one marker encodes a more specific value subset than the other marker (e.g. many possible moods / one specific mood)
    -   Feature Subset
        -   one marker encodes a larger featural subset than the other (e.g. pers + num vs. just num)
        -   contrast this with "partially superfluous multiple exponence" [@caballeroharris2012; @grofulovicmuller2023] → these are generally combinations of multiple exponence + cumulative exponence

### Predictions

-   upon encountering a compatible marker, speakers analyse the verb as belonging to the same domain

    -   i.e. constructional paradigm is salient
-   if an incompatible marker is encountered, a new domain is posited
-   construction-internal pauses are permitted if syntactic project is not finished

    -   i.e. construction is salient
    -   normally to do with expected edge-marking, i.e. right-edge marked supersets
-   prosodic afterthoughts can be integrated into a SVC

### Examples

::: {.callout-note collapse="true"}

#### [Meriam](https://glottolog.org/resource/languoid/id/meri1244)

(@meriam2) @piper1989, 100, example 3.65
```{r}
#| label: meriam
gloss_example(transliteration = "wi na-kasir weret-lare",
             glosses = "3NSG.SBJ 3NSG.SBJ-go_along AUX.PL-NPRS.PL",
             free_translation = "They were going along.")
```
:::

::: {.callout-note collapse="true"}

#### [Momu-Fas](https://glottolog.org/resource/languoid/id/fass1245)

(@momu1) @honeyman2016, 423, example 13.5a
```{r}
#| label: momu-1
gloss_example(transliteration = "Yefu=ne key munta pwenin ai yesy.",
             glosses = "husband:SG:POSS=FOC hand fold_arms\3SG.SUBJ stand[1|3SG.SUBJ] ANIM:there.be[1|3SG.SUBJ] just",
             free_translation = "As for her husband, there he was, just arms folded, standing.")
```

(@momu2) @honeyman2016, 427, example (b)
```{r}
#| label: momu-2
gloss_example(transliteration = "Ereta=pwe=Aitafe=m peeni.",
             glosses = "do.like.that=[one.]comes=Aitape=OBL [one.]arrives[1SG.SUBJ]",
             free_translation = "I came to Aitape like that.")
```
:::

::: {.callout-note collapse="true"}

#### [Numbami](https://glottolog.org/resource/languoid/id/numb1247)

Specifically, inclusory constructions:

(@numbami) @bradshaw1993, 149, example xxx (glossing adjusted)
```{r}
#| label: numbami
gloss_example(transliteration = "nu-tamu ai muna-wasa",
             glosses = "2SG.IRR-join 3PL 2PL-go",
             free_translation = "Join them and (you) go.")
```
:::

::: {.callout-note collapse="true"}

#### [Paamese](https://glottolog.org/resource/languoid/id/paam1238)

Specifically, inclusory constructions:

(@paamese4) @crowley1987, 47, example 23 (glossing adjusted)
```{r}
#| label: paamese-inclusory
gloss_example(transliteration = "ma-kuri-ko lo-va-haa",
             glosses = "1SG.IMM-take-2SG lDU.INCL-IMM-go",
             free_translation = "I will take you away with me.",
             annotation = "Makurik lovaha")
```
:::

::: {.callout-note collapse="true"}
#### [Zayse-Zergulla](https://glottolog.org/resource/languoid/id/zays1235)

> Non-final verbs in chains of two or more VP's having a common subject... tend to appear in the short perfect form. The overall aspect / tense setting of the sentence is determined by the form of the final verb. [@hayward1990a, 317-318]

(@zayse1) @hayward1990a, 318 (glossing adapted)
```{r}
#| label: zayse
gloss_example(transliteration = "ʔe=ʔas̀í gel-í ʔútt-i múud-i geháttesin",
             glosses = "DET.MASC=man.NOM enter-PFV sit-PFV eat-PFV sleep.3SG.MASC.PFV",
             free_translation = "The man returned home, sat down, ate, and went to sleep.")
```

:::

# Implications

some different paths for language change based on exponence pattern?

-   single exponence 

    → prosody is the pathway to language change

-   multiple exponence 
    
    → redundant markers due to information identity are likely to be lost or 'eroded' (cf. “optional” marking in Aikhenvald's sense)
-   subset exponence 

    → less specific markers more likely to be lost? [cf. @grofulovicmuller2023]
-   distributed exponence 

    → markers will be retained (maybe semantic shift will happen?)

# Conclusions

-   we can test these hypotheses

    -   and therefore refine our understanding of what a SVC is, especially wrt definitional characteristics

    →   the hypotheses here need to be refined! Multiple exponence already appears to be capturing more than one phenomenon at a time.
    
-   do you know relevant examples? or want to test these in your data/your fieldwork?

    -   get in touch via email (naomi.peck(at)linguistik.uni-freiburg.de)
    
# References